---
title: "meta testing (Reproducible)"
author: "salem malibary"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(pacman)
p_load(tidyverse, here, metafor, lubridate, RColorBrewer, gt, 
       labelled, writexl, ggpubr, stringr, readxl, robumeta, patchwork)

# Load real data
meta_data <- read.csv(here("data", "02 meta ready to analysis.csv"))
meta_data$variance <- (meta_data$standard_error)^2

# Global theme
my_theme <- theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    legend.box = "vertical",
    axis.text.y = element_text(family = "mono", size = 11)
  )
```

```{r eval=FALSE, include=FALSE}
labelled::var_label(meta_data)
```

```{r}
meta_data <- meta_data  |> 
  select(
    -Column1, -Column2, -Column3, -Column4, -Column5,
    -variance, -Row_ID, -Size..n., -significant,
    -author, -Name)
```

```{r}

```

```{r}
meta_data <- meta_data %>%
  mutate(
    n_total = case_when(
      !is.na(n_high) & !is.na(n_low) ~ n_high + n_low,
      !is.na(n_pre) & !is.na(n_post) ~ pmax(n_pre, n_post, na.rm = TRUE),
      !is.na(n_pre) ~ n_pre,
      !is.na(n_post) ~ n_post,
      !is.na(sample.size) ~ sample.size,
      TRUE ~ NA_real_
    )
  )
```

```{r}
meta_data <- meta_data %>%
  rename(
    group        = group,
    high         = group_high,
    low          = group_low,
    site         = site,
    study_id     = author_year,
    design       = study_design,
    m1           = m_high,
    m2           = m_low,
    sd1          = sd_high,
    sd2          = sd_low,
    smd          = smd_effctsize,
    se           = standard_error,
    ci_low       = lower_ci,
    ci_high      = upper_ci,
    t_stat       = t,
    df           = df,
    sd_pooled    = pooled_sd
  )
```

```{r}
meta_data <- meta_data %>%
  mutate(
    # SMD for between-group design
    smd_between = case_when(
      !is.na(m1) & !is.na(m2) & !is.na(sd1) & !is.na(sd2) & !is.na(n_high) & !is.na(n_low) ~ 
        (m1 - m2) / sqrt(((n_high - 1)*sd1^2 + (n_low - 1)*sd2^2) / (n_high + n_low - 2)),
      TRUE ~ NA_real_
    ),
    
    # SMD for within-group design
    smd_within = case_when(
      !is.na(m_post) & !is.na(m_pre) & !is.na(sd_pre) ~ 
        (m_post - m_pre) / sd_pre,
      !is.na(m_post) & !is.na(m_pre) & !is.na(sd_change) ~ 
        (m_post - m_pre) / sd_change,
      TRUE ~ NA_real_
    )
  )

```

```{r}
meta_data <- meta_data %>%
  mutate(
    # SE for between-group SMD
    se_between = case_when(
      !is.na(smd_between) & !is.na(n_high) & !is.na(n_low) ~ 
        sqrt((n_high + n_low) / (n_high * n_low) + (smd_between^2) / (2 * (n_high + n_low))),
      TRUE ~ NA_real_
    ),
    
    ci_low_between = smd_between - 1.96 * se_between,
    ci_high_between = smd_between + 1.96 * se_between,
    
    # SE for within-group SMD
    se_within = case_when(
      !is.na(smd_within) & !is.na(n_pre) ~ 
        sqrt(1 / n_pre + (smd_within^2) / (2 * n_pre)),
      TRUE ~ NA_real_
    ),
    
    ci_low_within = smd_within - 1.96 * se_within,
    ci_high_within = smd_within + 1.96 * se_within
  )

```

```{r}
meta_data <- meta_data %>%
  rename(
    id            = ID,
    group_high    = high,
    group_low     = low,
    n             = sample.size,
    m_high        = m1,
    m_low         = m2,
    sd_high       = sd1,
    sd_low        = sd2
  )

```

```{r}
names(meta_data) <- names(meta_data) %>%
  str_replace_all("\\.", "_") %>%
  str_replace_all("([a-z])([A-Z])", "\\1_\\2") %>%
  tolower()

```

```{r}
meta_data <- meta_data %>%
  select(-delete, -direction_description, -significant_direction)
```

```{r}
meta_data <- meta_data %>%
  mutate(
    smd_final = coalesce(smd_between, smd_within, smd),
    se_final  = coalesce(se_between, se_within, se)
  )
```

```{r}
comparison <- meta_data %>%
  select(study_id, smd, smd_final, se, se_final) %>%
  mutate(
    smd_diff = smd_final - smd,
    se_diff  = se_final - se
  )
```


```{r}
head(comparison, 10)  # or View(comparison)
```

```{r}
meta_data <- meta_data %>%
  mutate(
    use_final = case_when(
      !is.na(smd_final) & abs(smd_final - smd) < 0.2 ~ TRUE,
      is.na(smd) ~ TRUE,
      TRUE ~ FALSE
    )
  )

```

```{r}
meta_data %>%
  filter(use_final == FALSE)
```

```{r}
meta_data <- meta_data %>%
  mutate(
    smd_diff = smd_final - smd,
    smd_flip_flag = ifelse(sign(smd_final) != sign(smd) & !is.na(smd) & !is.na(smd_final), TRUE, FALSE)
  )

```

```{r}
meta_data %>% filter(smd_flip_flag == TRUE)

```

```{r}
meta_data <- meta_data %>%
  mutate(
    smd_final_harmonized = ifelse(smd_flip_flag == TRUE, -1 * smd_final, smd_final)
  )
```

```{r}
labelled::var_label(meta_data)$smd_final_harmonized <- "SMD (Harmonized Direction)"
```

```{r}
meta_model <- rma(yi = smd_final_harmonized, sei = se_final, data = meta_data, method = "REML")
summary(meta_model)
forest(meta_model, slab = meta_data$study_id)

```


```{r}
meta_data$exposure_time__mins_ <- as.numeric(meta_data$exposure_time__mins_)
```

```{r}
meta_model_time <- rma(
  yi = smd_final_harmonized,
  sei = se_final,
  mods = ~ exposure_time__mins_,
  data = meta_data,
  method = "REML"
)
summary(meta_model_time)


```

```{r eval=FALSE, include=FALSE}
bubble(meta_model_time, xlab = "Exposure Time (minutes)")
```

```{r}
meta_data$predicted <- predict(meta_model_time)$pred
```

----

meta regression 
```{r}
library(ggplot2)

ggplot(meta_data, aes(x = exposure_time__mins_, y = smd_final_harmonized)) +
  geom_point(aes(size = 1 / se_final^2), alpha = 0.6) +  # weight by precision
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Exposure Time (minutes)",
    y = "Harmonized SMD",
    title = "Meta-Regression: Effect Size by Exposure Time"
  ) +
  theme_minimal(base_size = 14)

```

The slope is very flat, meaning:
- No strong linear relationship between exposure time and effect size.
- The shaded confidence band is wide, especially at higher exposure times.
- This reflects uncertainty (fewer data points and more variance at longer durations).
- There's a strong outlier at very low exposure time with a huge negative SMD.
- That one study might heavily influence the model — worth checking separately.

----
```{r}
meta_data <- meta_data %>%
  mutate(
    exposure_time_group = case_when(
      is.na(exposure_time__mins_) ~ NA_character_,
      exposure_time__mins_ <= 10 ~ "short",
      exposure_time__mins_ <= 30 ~ "medium",
      TRUE ~ "long"
    ) %>% factor(levels = c("short", "medium", "long"))
  )
```

```{r}
meta_model_group <- rma(
  yi = smd_final_harmonized,
  sei = se_final,
  mods = ~ exposure_time_group,
  data = meta_data,
  method = "REML"
)
summary(meta_model_group)
forest(meta_model_group, slab = meta_data$study_id)
```

No significant effect of exposure time group:

p-value for the moderator: 0.9546 → not statistically significant

So, there's no evidence that SMD differs meaningfully across short, medium, or long exposure durations.

Extremely high heterogeneity:

I² = 99.65%

Tau² = 4.82 (very high)

This means the variation between studies is huge, and not explained by exposure_time_group.

Group estimates:

- "short" (intercept): -0.52 (not significant)

- "medium": +0.17 higher than short → not significant

- "long": +0.50 higher than short → also not 

----

```{r}
funnel(meta_model)
```

```{r}
tf_model <- trimfill(meta_model)
funnel(tf_model, main = "Funnel Plot with Trim-and-Fill")
```

There’s a clear clustering of studies on the right side (positive SMDs), and a gap on the left.
The trim-and-fill added several white circles on the left — suggesting potential publication bias, where small or negative-effect studies might be missing.
overall effect may be inflated due to missing studies with smaller or negative results.
After correction, the adjusted effect size is lower (closer to zero).

```{r}
regtest(meta_model, model = "rma", predictor = "sei")
```

----

```{r}
table(meta_data$design)
```

```{r}
meta_data <- meta_data %>%
  mutate(
    design_group = case_when(
      str_detect(tolower(design), "within") ~ "within",
      str_detect(tolower(design), "pre") ~ "pre_post",
      str_detect(tolower(design), "between") ~ "between",
      TRUE ~ "other"
    ) %>% factor(levels = c("between", "pre_post", "within", "other"))
  )
```

```{r}
meta_model_clean <- rma(
  yi = smd_final_harmonized,
  sei = se_final,
  mods = ~ design_group,
  data = meta_data,
  method = "REML"
)
summary(meta_model_clean)

```

- within-subject designs tend to produce lower SMDs than between designs
- The p-value is close to 0.05 → potentially meaningful, but not statistically confirmed
- pre_post is not meaningfully different from between

----
```{r}
meta_data <- meta_data %>%
  mutate(
    assessment_group = case_when(
      str_detect(tolower(outcome_assessment), "pss") ~ "pss",
      str_detect(tolower(outcome_assessment), "dass") ~ "dass",
      str_detect(tolower(outcome_assessment), "tmd") ~ "tmd",
      str_detect(tolower(outcome_assessment), "eeg|alpha") ~ "eeg",
      str_detect(tolower(outcome_assessment), "conductance|scl|scr") ~ "sc",
      TRUE ~ "other"
    ) %>% factor(levels = c("pss", "dass", "tmd", "eeg", "sc", "other"))
  )

meta_model_assess <- rma(
  yi = smd_final_harmonized,
  sei = se_final,
  mods = ~ assessment_group,
  data = meta_data,
  method = "REML"
)

```

```{r}
coefs <- summary(meta_model_assess)$b
ci_low <- summary(meta_model_assess)$ci.lb
ci_high <- summary(meta_model_assess)$ci.ub
```

```{r}
coef_df <- data.frame(
  group = str_replace(rownames(coefs), "assessment_group", ""),
  estimate = coefs[, 1],
  ci_low = ci_low,
  ci_high = ci_high
)
```

```{r}
# Create the assessment_group variable
meta_data <- meta_data %>%
  mutate(
    assessment_group = case_when(
      str_detect(tolower(outcome_assessment), "pss") ~ "pss",
      str_detect(tolower(outcome_assessment), "dass") ~ "dass",
      str_detect(tolower(outcome_assessment), "tmd") ~ "tmd",
      str_detect(tolower(outcome_assessment), "eeg|alpha") ~ "eeg",
      str_detect(tolower(outcome_assessment), "conductance|scl|scr") ~ "sc",
      TRUE ~ "other"
    ) %>% factor(levels = c("pss", "dass", "tmd", "eeg", "sc", "other"))
  )
```

```{r}
meta_model_assess <- rma(
  yi = smd_final_harmonized,
  sei = se_final,
  mods = ~ assessment_group,
  data = meta_data,
  method = "REML"
)
```

```{r}
# Clean model output
coefs <- summary(meta_model_assess)$b
ci_low <- summary(meta_model_assess)$ci.lb
ci_high <- summary(meta_model_assess)$ci.ub

# Fix group names and create base dataframe
coef_df <- data.frame(
  group = str_replace(rownames(coefs), "assessment_group", ""),
  estimate = coefs[, 1],
  ci_low = ci_low,
  ci_high = ci_high
)

# Study counts
group_counts <- meta_data %>%
  group_by(assessment_group) %>%
  summarise(k = n(), .groups = "drop") %>%
  mutate(assessment_group = as.character(assessment_group))

# Merge
coef_df <- left_join(coef_df, group_counts, by = c("group" = "assessment_group"))

# Labeling
coef_df$group_label <- paste0(coef_df$group, " (n = ", coef_df$k, ")")
coef_df$label <- paste0(
  "SMD = ", sprintf("%.2f", coef_df$estimate), "\n",
  "[", sprintf("%.2f", coef_df$ci_low), ", ", sprintf("%.2f", coef_df$ci_high), "]\n",
  "n = ", coef_df$k
)
coef_df$significant <- ifelse(coef_df$ci_low > 0 | coef_df$ci_high < 0, "yes", "no")

# Reorder for plotting
coef_df <- coef_df %>%
  arrange(estimate) %>%
  mutate(group_label = factor(group_label, levels = group_label))

# Plot
ggplot(coef_df, aes(x = estimate, y = group_label, color = significant)) +
  # Cohen's d bands
  geom_rect(data = data.frame(), aes(xmin = -0.2, xmax = 0.2, ymin = -Inf, ymax = Inf),
            fill = "gray90", alpha = 0.4, inherit.aes = FALSE) +
  geom_rect(data = data.frame(), aes(xmin = -0.5, xmax = -0.2, ymin = -Inf, ymax = Inf),
            fill = "gray85", alpha = 0.3, inherit.aes = FALSE) +
  geom_rect(data = data.frame(), aes(xmin = 0.2, xmax = 0.5, ymin = -Inf, ymax = Inf),
            fill = "gray85", alpha = 0.3, inherit.aes = FALSE) +
  geom_rect(data = data.frame(), aes(xmin = -0.8, xmax = -0.5, ymin = -Inf, ymax = Inf),
            fill = "gray80", alpha = 0.2, inherit.aes = FALSE) +
  geom_rect(data = data.frame(), aes(xmin = 0.5, xmax = 0.8, ymin = -Inf, ymax = Inf),
            fill = "gray80", alpha = 0.2, inherit.aes = FALSE) +
  geom_rect(data = data.frame(), aes(xmin = -Inf, xmax = -0.8, ymin = -Inf, ymax = Inf),
            fill = "gray75", alpha = 0.2, inherit.aes = FALSE) +
  geom_rect(data = data.frame(), aes(xmin = 0.8, xmax = Inf, ymin = -Inf, ymax = Inf),
            fill = "gray75", alpha = 0.2, inherit.aes = FALSE) +

  # Labels for zones
  annotate("text", x = 0, y = Inf, label = "Negligible", vjust = -1, size = 3.5, color = "gray40") +
  annotate("text", x = 0.35, y = Inf, label = "Small", vjust = -1, size = 3.5, color = "gray40") +
  annotate("text", x = 0.65, y = Inf, label = "Medium", vjust = -1, size = 3.5, color = "gray40") +
  annotate("text", x = 1.2, y = Inf, label = "Large", vjust = -1, size = 3.5, color = "gray40") +
  annotate("text", x = -0.35, y = Inf, label = "Small", vjust = -1, size = 3.5, color = "gray40") +
  annotate("text", x = -0.65, y = Inf, label = "Medium", vjust = -1, size = 3.5, color = "gray40") +
  annotate("text", x = -1.2, y = Inf, label = "Large", vjust = -1, size = 3.5, color = "gray40") +

  # Main plot
  geom_point(size = 4) 


```

